
# R Setup

```{r}
library(here)
library(tidyverse)
library(fs)
# library(speedyseq)
# library(metacal)

# library(cowplot)
# theme_set(theme_cowplot())
# library(patchwork)
# library(ggbeeswarm)

import::from(Biostrings, readDNAStringSet, DNAString, DNAStringSet,
  reverseComplement)
```

# Sequence setup

Read in the sequences output by barCoder,


```{r}
base_dir <- here('analysis/barcoder-projects/2022-05-05-barcode-primers')
# fns <- c(barcode = 'barcodeList.fa', primer = 'primerList.fa') %>%
barcodes <- path(base_dir, 'barcodeList.fa') %>%
  readDNAStringSet()
primers <- path(base_dir, 'primerList.fa') %>%
  readDNAStringSet()
dna <- c(barcodes, primers)
```

Parse the sequence names to get a table with info about each sequence.

```{r}
dna_tbl <- dna %>%
  as.character %>%
  enframe(value = 'sequence') %>%
  separate(name, into = c('target_genome', 'sequence_type', 'replicate'),
    remove = FALSE, sep = ', ') %>%
  mutate(
    across(target_genome, str_replace, '^[ ]+', ''),
    sequence_type = ifelse(sequence_type == '', 'Barcode', sequence_type),
    across(replicate, str_extract, '[0-9]+$'),
    across(replicate, as.integer),
    length = nchar(sequence)
  ) %>%
  glimpse
```

# Get the spacer and probe start and stop locations

Compute the spacer locations from the primer and barcode lengths,

```{r}
dna_pos <- dna_tbl %>%
  select(target_genome, replicate, sequence_type, length) %>%
  pivot_wider(names_from = sequence_type, values_from = length,
    names_glue = '{.name}_length'
  ) %>%
  janitor::clean_names() %>%
  mutate(
    spacer_start = target_primer_1_length + 1,
    spacer_end = barcode_length - target_primer_2_length,
  )
```

Find the probe locations using string matching

```{r}
probe_pos <- dna_tbl %>%
  select(target_genome, replicate, sequence_type, sequence) %>%
  pivot_wider(names_from = sequence_type, values_from = sequence) %>%
  janitor::clean_names() %>%
  mutate(
    probe_loc = str_locate(barcode, target_probe),
    across(probe_loc, as_tibble),
  ) %>%
  unpack(probe_loc, names_sep = '_') %>%
  rename_with(.cols = starts_with('probe_loc'), str_replace, '_loc', '') %>%
  glimpse

#> check
tmp <- probe_pos %>%
  transmute(target_probe, 
    pred = str_sub(barcode, probe_start, probe_end)
  )
stopifnot(all.equal(tmp$target_probe, tmp$pred))
rm(tmp)
```

Create a wide table with all sequences and positions,

```{r}
dna_wide <- left_join(
  probe_pos, 
  dna_pos, 
  by = c("target_genome", "replicate")
) %>%
  mutate(
    spacer = str_sub(barcode, spacer_start, spacer_end),
  # NOTE: Need to reverse complement the primer seq to construct barcode
    target_primer_2_rc = map_chr(target_primer_2,
      ~ DNAString(.x) %>% reverseComplement %>% as.character)
  )
# TODO: check that adding the primers + spacer gives the original barcode
```

# Create final barcodes (keeping original probes)

Use the first barcode as the model to get left and right spacer sequences from,

```{r}
model <- dna_wide %>%
  filter(replicate == 1) %>%
  mutate(
    spacer_1 = str_sub(barcode, spacer_start, probe_start - 1),
    spacer_2 = str_sub(barcode, probe_end + 1, spacer_end)
  )
  
# check
s1 <- model$barcode
s2 <- str_c(
  model$target_primer_1, 
  model$spacer_1, 
  model$target_probe, 
  model$spacer_2, 
  model$target_primer_2_rc
  #> model$target_primer_2 %>% DNAString %>% reverseComplement %>% as.character
)
stopifnot(identical(s1, s2))
```

```{r}
model$spacer_1
#> [1] "GTGCCCTGTAATCTATCTGTCAGTCAAGC"
model$spacer_2
#> [1] "AGGTCCACTCCCCTCGTCACTATACCAAGAA"
```

Check the length distribution of the original barcodes,

```{r}
dna_wide %>% pull(barcode_length) %>% table
#> .
#> 132 133 134 135 136 
#>  13  12  17   4   4 
```

We can remove up to 7 bp from the spacers.

```{r}
model <- model %>%
  mutate(across(c(spacer_1, spacer_2), str_sub, end = -4))
model$spacer_1
#> [1] "GTGCCCTGTAATCTATCTGTCAGTCA"
model$spacer_2
#> [1] "AGGTCCACTCCCCTCGTCACTATACCAA"
```

```{r}
dna_wide1 <- dna_wide %>%
  mutate(
    new_barcode = str_c(target_primer_1, model$spacer_1, target_probe,
      model$spacer_2, target_primer_2_rc),
  )
```

For the final export, might be nice to have updated lengths and start/stop positions,

```{r}
final_tbl <- dna_wide1 %>%
  transmute(
    target_genome, replicate,
    barcode = new_barcode,
    barcode_length = nchar(barcode),
    primer_1 = target_primer_1, 
    primer_1_length = nchar(primer_1), 
    primer_2 = target_primer_2, 
    primer_2_length = nchar(primer_2), 
    probe = target_probe,
    probe_length = nchar(probe),
  )
```

Let's also save a version of this table with the original barcodes,

```{r}
org_tbl <- dna_wide %>%
  transmute(
    target_genome, replicate,
    barcode,
    barcode_length = nchar(barcode),
    primer_1 = target_primer_1, 
    primer_1_length = nchar(primer_1), 
    primer_2 = target_primer_2, 
    primer_2_length = nchar(primer_2), 
    probe = target_probe,
    probe_length = nchar(probe),
  )
```

## Save output

```{r}
# dir_create(here('output'))
dir_create(here('output/2022-05-05-barcode-primers'))
```

```{r}
final_tbl %>%
  write_csv(here('output/2022-05-05-barcode-primers', 'barcodes-postprocessed.csv'))
org_tbl %>%
  write_csv(here('output/2022-05-05-barcode-primers', 'barcodes-unmodified.csv'))
```

```{r}
org_tbl %>% pull(barcode_length) %>% table
#> .
#> 132 133 134 135 136 
#>  13  12  17   4   4 
final_tbl %>% pull(barcode_length) %>% table
#> .
#> 126 127 128 129 130 
#>  13  12  17   4   4 
```

