

```{r}
library(tidyverse)
library(here)
library(generics)

library(cowplot)
theme_set(theme_cowplot())
```

```{r}
# concentration in copies per microliter
conc_low <- 0.1
dilution_step <- 20
num_samples <- 7
conc_max <- conc_low * dilution_step^(num_samples - 1)

dilution_df <- tibble(
  dilution_power = 0:6,
  dilution_factor = dilution_step^dilution_power,
  conc = conc_max / dilution_factor,
  dilution_name = str_c('D', '0', dilution_power + 1)
)
```

## Explore single qPCR experiment, Cq values


```{r}
x <- 'data/2022-06-29-spike-in-experiment/results/2022-06-29-trip01.xlsx' %>%
  here() %>%
  readxl::read_excel(sheet = 'Results', skip = 40,
    na = c("Undetermined")
  ) %>%
  janitor::clean_names() %>%
  glimpse
  #> ReadqPCR::read.taqman()
```

```{r}
x1 <- x %>%
  separate(sample_name, 
    into = c('matrix', 'target_name', 'dilution_name'),
    sep = '_', remove = FALSE
  ) %>%
  left_join(dilution_df, by = 'dilution_name')
```

```{r}
x1 %>%
  ggplot(aes(conc, ct, color = target_name)) +
  scale_x_log10() +
  geom_point(size = 2)
```


Estimate the efficiency from the slope of the standard curves:

```{r}
x1 %>%
  #> filter(!is.na(target_name), conc > 200) %>%
  filter(!is.na(target_name), conc > 1e4) %>%
  with_groups(target_name, nest) %>%
  mutate(
    fit = map(data, ~lm(ct ~ log10(conc), data = .x)),
    fit = map(fit, tidy)
  ) %>%
  unnest(fit) %>%
  filter(term == 'log10(conc)') %>%
  transmute(target_name, slope = estimate,
    efficiency_hat = 10^(-1 / slope) - 1
  )
```

Note, we get a better E-hat (closer to 1) when weuse the middle concentration, but I suspect that that is spurious - I suspect that the curve is flattening out there (not linear anymore) and it's causing the slope estimate to flatten out, which increases the estimated efficiency.

Next steps involve 

- determining the linear range, and estimating the efficiency from this
- checking the NTC
- adding the 'background' levels to the plot, like Anjali did
<!--  -->

## Read in all qPCR experiments

## Explore amplification data

- what do the trajectories look like, in the plateau zone?
- can I reproduce the ct calculations?
  - can I reproduce the delta-rn calculations? this would involve computing Rn-; see https://www.thermofisher.com/us/en/home/life-science/pcr/real-time-pcr/real-time-pcr-learning-center/real-time-pcr-basics/essentials-real-time-pcr.html and  https://www.mhtpmedicalgenomics.org.au/images/service-pdfs/Understanding_curves.pdf
<!--  -->

```{r}
sam <- x1 %>%
  select(starts_with('well'), sample_name, matrix, target_name, 
    starts_with('dilution'), conc,
    starts_with('baseline') # for computing and vis of baselines
  ) %>%
  distinct
```

TODO: Rethink through 'sample_name' and 'target_name'; note that 'target name' in the spreadsheet is not the same as what I define it as above (which is instead the primer/probe set)


```{r}
amp <- 'data/2022-06-29-spike-in-experiment/results/2022-06-29-trip01.xlsx' %>%
  here() %>%
  readxl::read_excel(sheet = 'Amplification Data', skip = 40,
    na = c("Undetermined"),
    col_types = c('numeric', 'text', 'numeric', 'text', 'numeric', 'numeric')
  ) %>%
  janitor::clean_names() %>%
  select(-target_name) %>%
  left_join(sam, by = c('well', 'well_position')) %>%
  glimpse
```

```{r}
amp %>%
  filter(!is.na(target_name)) %>%
  ggplot(aes(cycle, rn, color = target_name)) +
  facet_wrap(~dilution_name) +
  geom_line(aes(group = well_position)) +
  geom_point()
```

Why do the baselines differ for the two targets? Does this have to do with plate geometry, or is it a feature of the probe?


```{r}
amp %>%
  filter(!is.na(target_name)) %>%
  ggplot(aes(cycle, delta_rn, color = target_name)) +
  facet_wrap(~dilution_name) +
  geom_line(aes(group = well_position)) +
  geom_point()
```

```{r}
amp %>%
  filter(!is.na(target_name)) %>%
  ggplot(aes(cycle, delta_rn, color = target_name)) +
  facet_wrap(~dilution_name) +
  #> scale_y_log10(limits = c(1e-3, 1e1)) +
  scale_y_log10() +
  geom_line(aes(group = well_position)) +
  geom_point()
```

### understanding baselines

To compute Delta-Rn from Rn, can compute the Rn- from the reported baselines from the software,

Perhaps would be good to start by doing this by hand for a single case?

```{r}
baselines <- amp %>%
  filter(cycle >= baseline_start & cycle <= baseline_end) %>%
  with_groups(well, summarize,
    rn_baseline = mean(rn)
  )
amp1 <- amp %>%
  left_join(baselines, by = 'well') %>%
  mutate(delta_rn_new = rn - rn_baseline)
```

```{r}
amp1 %>%
  select(starts_with('delta_rn'))
```


what if I look at the difference bwetween delta_rn and rn from the software output?


```{r}
amp %>%
  filter(!is.na(target_name)) %>%
  ggplot(aes(cycle, rn - delta_rn, color = target_name)) +
  facet_wrap(~dilution_name) +
  #> scale_y_log10(limits = c(1e-3, 1e1)) +
  #> scale_y_log10() +
  geom_line(aes(group = well_position)) +
  geom_point()
```

Why is this not constant? According to the below formula, it seems like Rn - DeltaRn should equal a constant baseline value.

Perhaps what the software does is fit a line, so that the baseline is actually not constant but a line with a non-zero slope.


# Notes

Next steps are going to involve debugging why we are getting amplification 

Should plot the plate design, to make sure I understand, and consider the source of the Rn-.


I don't understand how Delta Rn is calculated. 
From the software manual,

> •  The data collected is the normalized intensity of the reporter dye, or Rn.
•  If end-point experiments include pre-PCR data points, the software calculates the delta Rn (ΔRn) value according to the following formula:
ΔRn = Rn (post-PCR read) – Rn (pre-PCR read), where Rn = normalized readings.

However, the difference between delta_rn and rn seems to differ by cycle.
