
# R Setup

```{r}
library(here)
library(tidyverse)
library(fs)
library(googlesheets4)
# library(speedyseq)
# library(metacal)

# library(cowplot)
# theme_set(theme_cowplot())
# library(patchwork)
# library(ggbeeswarm)

import::from(Biostrings, readDNAStringSet, DNAString, DNAStringSet,
  writeXStringSet, reverseComplement)
```

# Sequence setup

Read in the primer and barcode sequences in the first and second sets,

```{r}
fns <- c(
  'output/2022-05-02-barcode-primers/barcodes-postprocessed.csv',
  'output/2022-05-05-barcode-primers/barcodes-postprocessed.csv'
) %>%
  here
x <- tibble(set = 1:2, file = fns) %>%
  mutate(
    data = map(file, read_csv, col_types = 'cicicicici')
  ) %>%
  unnest(data) %>%
  glimpse
```

We can subset to just the barcodes that were ordered, as these are the only viable candidates we need to evaluate.

For set 1, Anjali ordered all barcodes >= 125bp in length.
For set 2, Anjali ordered 20 barcodes passing a secondary-structure check; we can find this set from a Google Sheet.

```{r}
set2_sheet <- 
  # 'https://docs.google.com/spreadsheets/d/1vx3hgAowibfDIi7tG7E5eALsZu7wZEX22ZD4ph5slw4/edit#gid=639702623'
'https://docs.google.com/spreadsheets/d/1vx3hgAowibfDIi7tG7E5eALsZu7wZEX22ZD4ph5slw4'
ag_thermo <- read_sheet(set2_sheet, 'Thermo Probes')
```

Confirm compatibility of the two sheets:

```{r}
ag1 <- ag_thermo %>%
  janitor::clean_names() %>%
  select(assay_name, 
    primer_1 = forward_primer_sequence, 
    primer_2 = reverse_primer_sequence, 
    probe = probe_sequence)
x1 <- x %>%
  mutate(
    across(replicate, formatC, width = 3, flag = "0"),
    assay_name = str_glue('qpcr_backbone_s{set}_{replicate}') %>% as.character
  )
x2 <- x1 %>%
  filter(assay_name %in% ag1$assay_name) %>%
  select(assay_name, primer_1, primer_2, probe)
stopifnot(all.equal(x2, ag1))
```

Let's create a new dataframe marking which were purchased and creating unique barcode names,

```{r}
barcodes <- x1 %>%
  mutate(
    purchased = (set == 1 & barcode_length >= 125) | 
      (set == 2 & assay_name %in% ag1$assay_name)
  )
```

Next, let's create a DNAStringSet of the primers and probes, with unique identifiers.
(Subsetting to just the purchased ones)
I'll make the sequence ids of the format `{assay_name}_{sequence type}`.

```{r}
primers <- barcodes %>%
  filter(purchased) %>%
  select(assay_name, primer_1, primer_2, probe) %>%
  pivot_longer(-assay_name, names_to = 'type', values_to = 'sequence') %>%
  mutate(
    across(type, str_replace, '_', ''), # Remove underscores
    sequence_id = str_glue('{assay_name}_{type}')
  ) %>%
  glimpse

primers_dna <- primers %>%
  select(sequence_id, sequence) %>%
  deframe %>%
  DNAStringSet
```

Export Fasta file for providing to blast,

```{r}
dir_create(here('output', '2022-05-12-blast'))
query_fn <- here('output', '2022-05-12-blast', 'primers-and-probes.fasta')
writeXStringSet(primers_dna, query_fn)
```

# Blast against nt

This was done on a cluster; see submit.sh.

# Blast against each other

This we can do locally,

```sh
makeblastdb -in primers-and-probes.fasta -parse_seqids -title primers-and-probes -dbtype nucl
mkdir blastdb
mv primers-and-probes.fasta.* blastdb
# blast all against all
n=$(nproc)
blastn -db ./blastdb/primers-and-probes.fasta -query primers-and-probes.fasta \
    -task blastn-short \
    -num_threads $n \
    -out all-against-all.tsv -outfmt '6 std qlen score'
gzip all-against-all.tsv
```


