

This file re-runs the colony picking simulation with our new phagemid barcode design, in wich we have 5 qPCR barcodes, each receiving ~20 sequencing barcodes (96 total), and are aiming to get 5X5.

```{r}
library(here)
library(tidyverse)
library(ggridges)
library(cowplot)

library(furrr)
plan('multicore', workers = 3) 
```

Generate the ids for the library,
We have 96 total sequencing barcodes, and I will distribute these evenly over the 5 qPCR barcodes, so that qPCR id 1 has 20 seq ids and qPCR id 2:5 have 19 seq ids.

```{r}
barcodes <- crossing(qpcr_id = 1:5, seq_subid = 1:20) %>%
  filter(!(seq_subid == 10 & qpcr_id >= 2)) %>%
  mutate(
    # across(ends_with('id'), as.character),
    seq_id = str_glue('{qpcr_id}_{seq_subid}') %>% as.character,
  ) %>%
  glimpse
```

# Functions

```{r}
# Comipute the min number of unique seq ids per qpcr id
min_seq <- function(sim) {
  sim %>%
    with_groups(qpcr_id, summarize, num_seq_ids = unique(seq_id) %>% length) %>%
    pull(num_seq_ids) %>% 
    min
}
#> min_seq(sim %>% slice_head(n=100))

# compute the cumulative number of unique seq ids
cum_uniq <- function(sim) {
  map_int(seq(nrow(sim)), 
    ~ sim$seq_id %>% head(.x) %>% unique(.) %>% length
  )
}

# compute the cumulative number of min unique seq ids per qpcr id
cum_min_seq <- function(sim) {
  map_int(seq(nrow(sim)), 
    ~ sim %>% head(n = .x) %>% min_seq
  )
}
```

## Actual simulations

For these, I think we can get away with a much smaller max colony number,

```{r, eval = F}
set.seed(42)
num_sims <- 500
t_max <- 200

sims <- tibble(.id = seq(num_sims)) %>%
  mutate(
    # Simulate picking t_max colonies
    sim = map(.id, ~slice_sample(barcodes, n = t_max, replace = TRUE)),
    # Compute the cumulative numbers of unique seq ids
    sim_uniq = future_map(sim, cum_uniq),
    # Compute the cumulative numers of min seq ids / qpcr id
    sim_min_seq = future_map(sim, cum_min_seq)
  )

saveRDS(sims, '2022-06-01-simulations.Rds')
#> sims <- readRDS('simulations.Rds')
```


```{r}
# Number of unique seq barcodes (condition 1)
M <- 25
# Number of unique seq barcodes / qpcr barcode (condition 2)
m <- 5

sims_first <- sims %>%
  transmute(.id,
    first_uniq = map_int(sim_uniq, ~which(. == M)[1]),
    first_min_seq = map_int(sim_min_seq, ~which(. == m)[1]),
  )
# Check that the colony numbers are always less than t_max
sims_first %>% summarize(across(starts_with('first'), max))
```

check the distributions,

```{r}
sims_first_long <- sims_first %>%
  pivot_longer(-.id, names_to = 'criterion') %>%
  mutate(across(criterion, fct_recode,
    `25 total` = 'first_uniq',
    `5 per qPCR id` = 'first_min_seq',
    )
  )
```


```{r}
sims_first_long %>%
  ggplot(aes(x = value, y = criterion, height = ..density..)) +
  geom_density_ridges(
    stat = 'density'
    # jittered_points = TRUE
  ) +
  labs(
    y = 'Criterion',
    x = 'Number of colonies'
  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(50, 100, 150, 200)) +
  scale_y_discrete(expand = c(0, 0)) +
  #> coord_cartesian(clip = 'off') +
#> scale_x_continuous(expand = c(0, 0))+
  expand_limits(x = c(50, 200)) +
  theme_ridges()
ggsave('/tmp/sims.png', width = 5, height = 3)
```

Can see that we need to pick many more colonies if we want at least 5 unique sequence ids for each qPCR id.


```{r}
sims_first_long %>%
  ggplot(aes(x = value)) +
  stat_ecdf(aes(color = criterion), size = 0.8) +
  geom_hline(yintercept = 0.9, linetype = 'dashed') +
  labs(
    color = 'Criterion',
    x = 'Number of colonies',
    y = 'Probability'
  ) +
  scale_x_continuous(breaks = c(50, 100, 150, 200)) +
  # scale_y_discrete(expand = c(0, 0)) +
  expand_limits(x = c(50, 200)) +
  theme_ridges() +
  theme(legend.position = 'top') +
  scale_color_brewer(type = 'qual')

ggsave('/tmp/sims-ecdf.png', width = 5, height = 3)
```

```{r}
p <- c(0.5, 0.8, 0.9, 0.95)
sims_first_long %>%
  with_groups(criterion, summarize, 
    across(value, quantile, probs = p), probability = p
  ) %>%
  knitr::kable()

```

|criterion     | value| probability|
|:-------------|-----:|-----------:|
|5 per qPCR id |  42.0|        0.50|
|5 per qPCR id |  50.2|        0.80|
|5 per qPCR id |  55.0|        0.90|
|5 per qPCR id |  60.0|        0.95|
|25 total      |  29.0|        0.50|
|25 total      |  30.0|        0.80|
|25 total      |  32.0|        0.90|
|25 total      |  33.0|        0.95|
