
# R Setup

```{r}
library(here)
library(tidyverse)
library(fs)
# library(speedyseq)
# library(metacal)

# library(cowplot)
# theme_set(theme_cowplot())
# library(patchwork)
# library(ggbeeswarm)

import::from(Biostrings, readDNAStringSet, DNAString, DNAStringSet,
  reverseComplement)
```

# Sequence setup

Read in the sequences output by barCoder,


```{r}
base_dir <- here('analysis/barcoder-projects/2022-05-02-barcode-primers')
# fns <- c(barcode = 'barcodeList.fa', primer = 'primerList.fa') %>%
barcodes <- path(base_dir, 'barcodeList.fa') %>%
  readDNAStringSet()
primers <- path(base_dir, 'primerList.fa') %>%
  readDNAStringSet()
dna <- c(barcodes, primers)
```

Parse the sequence names to get a table with info about each sequence.

```{r}
dna_tbl <- dna %>%
  as.character %>%
  enframe(value = 'sequence') %>%
  separate(name, into = c('target_genome', 'sequence_type', 'replicate'),
    remove = FALSE, sep = ', ') %>%
  mutate(
    across(target_genome, str_replace, '^[ ]+', ''),
    sequence_type = ifelse(sequence_type == '', 'Barcode', sequence_type),
    across(replicate, str_extract, '[0-9]+$'),
    across(replicate, as.integer),
    length = nchar(sequence)
  ) %>%
  glimpse
```

# Get the spacer and probe start and stop locations

Compute the spacer locations from the primer and barcode lengths,

```{r}
dna_pos <- dna_tbl %>%
  select(target_genome, replicate, sequence_type, length) %>%
  pivot_wider(names_from = sequence_type, values_from = length,
    names_glue = '{.name}_length'
  ) %>%
  janitor::clean_names() %>%
  mutate(
    spacer_start = target_primer_1_length + 1,
    spacer_end = barcode_length - target_primer_2_length,
  )
```

Find the probe locations using string matching

```{r}
probe_pos <- dna_tbl %>%
  select(target_genome, replicate, sequence_type, sequence) %>%
  pivot_wider(names_from = sequence_type, values_from = sequence) %>%
  janitor::clean_names() %>%
  mutate(
    probe_loc = str_locate(barcode, target_probe),
    across(probe_loc, as_tibble),
  ) %>%
  unpack(probe_loc, names_sep = '_') %>%
  rename_with(.cols = starts_with('probe_loc'), str_replace, '_loc', '') %>%
  glimpse

#> check
tmp <- probe_pos %>%
  transmute(target_probe, 
    pred = str_sub(barcode, probe_start, probe_end)
  )
stopifnot(all.equal(tmp$target_probe, tmp$pred))
rm(tmp)
```

Create a wide table with all sequences and positions,

```{r}
dna_wide <- left_join(
  probe_pos, 
  dna_pos, 
  by = c("target_genome", "replicate")
) %>%
  mutate(
    spacer = str_sub(barcode, spacer_start, spacer_end),
  # NOTE: Need to reverse complement the primer seq to construct barcode
    target_primer_2_rc = map_chr(target_primer_2,
      ~ DNAString(.x) %>% reverseComplement %>% as.character)
  )
# TODO: check that adding the primers + spacer gives the original barcode
```

# Create final barcodes (keeping original probes)

Use the first barcode as the model to get left and right spacer sequences from,

```{r}
model <- dna_wide %>%
  filter(replicate == 1) %>%
  mutate(
    spacer_1 = str_sub(barcode, spacer_start, probe_start - 1),
    spacer_2 = str_sub(barcode, probe_end + 1, spacer_end)
  )
  
# check
s1 <- model$barcode
s2 <- str_c(
  model$target_primer_1, 
  model$spacer_1, 
  model$target_probe, 
  model$spacer_2, 
  model$target_primer_2_rc
  #> model$target_primer_2 %>% DNAString %>% reverseComplement %>% as.character
)
stopifnot(identical(s1, s2))
```

```{r}
model$spacer_1
#> [1] "AAAATTTTAATTATTAAAAAATATAA"
model$spacer_2
#> [1] "TATAATAATTAATATTATATATATAA"
```

The spacers outside the probe are all A/T; not sure why.

For now, let's just use random spacers.

```{r}
set.seed(43)
model$spacer_1 <- sample(c('A', 'C', 'G', 'T'), nchar(model$spacer_1), replace = TRUE) %>% str_c(collapse = '')
model$spacer_2 <- sample(c('A', 'C', 'G', 'T'), nchar(model$spacer_2), replace = TRUE) %>% str_c(collapse = '')
model$spacer_1
#> [1] "TTTACACTTAGGCGCGCATGAATGAC"
model$spacer_2
#> [1] "TCTAGCGAAACTTCTTTGGAACACAG"
```

Create new barcodes where the left and right spacer regions are set to that of the model (first) barcode.
Note, the spacer and barcode lengths will be changed from the originals.

```{r}
dna_wide1 <- dna_wide %>%
  mutate(
    new_barcode = str_c(target_primer_1, model$spacer_1, target_probe,
      model$spacer_2, target_primer_2_rc),
  )
```

For the final export, might be nice to have updated lengths and start/stop positions,

```{r}
final_tbl <- dna_wide1 %>%
  transmute(
    target_genome, replicate,
    barcode = new_barcode,
    barcode_length = nchar(barcode),
    primer_1 = target_primer_1, 
    primer_1_length = nchar(primer_1), 
    # primer_1_start = 1,
    # primer_1_end = primer_1_start + primer_1_length,
    primer_2 = target_primer_2, 
    primer_2_length = nchar(primer_2), 
    # primer_2_start = ,
    # primer_2_end = ,
    probe = target_probe,
    probe_length = nchar(probe),
  )
```


## Save output


```{r}
dir_create(here('output'))
dir_create(here('output/2022-05-02-barcode-primers'))
```

```{r}
final_tbl %>%
  write_csv(here('output/2022-05-02-barcode-primers', 'barcodes-postprocessed.csv'))
```


# Create final barcodes (without probes)

Another option; should be easy based on the above, should we wish to ditch the probes.


