


```{r}
library(tidyverse)
library(here)
```

Hard constraints:
- Each tracer will appear in each of the 7 dilutions in exactly one sample.
<!--  -->

Other aims
- Each tracer should experience a wide range of proportions
- Less important but also desirable, is that pairwise ratios should have good variation.
- We are undecided on importance or desirability of total concentration variation
<!--  -->

We need to generate a table (matrix) to share and for Anjali to print out.
Could have a graphical and CSV version.
The table will have samples as rows, tracers as columns.

let's start by doing random permutations

```{r}
# concentration in copies per microliter
conc_low <- 0.1
dilution_step <- 20
num_samples <- 7
conc_max <- conc_low * dilution_step^(num_samples - 1)

dilution_df <- tibble(
  dilution_power = 0:6,
  dilution_factor = dilution_step^dilution_power,
  conc = conc_max / dilution_factor,
  dilution_name = str_c('D', dilution_power + 1)
)
dilution_df %>% knitr::kable()
```

Note, using Brian's scheme where D1 is undiluted / max concentration.


```{r}
#> vals <- str_c('D', 1:7)
vals <- dilution_df$dilution_name

set.seed(42)

mat <- map(seq(num_tracers), ~sample(vals)) %>% 
  unlist %>%
  matrix(ncol = num_tracers, byrow = FALSE)
rownames(mat) <- str_c('sample_', seq(num_samples))
colnames(mat) <- str_c('tracer_', seq(num_tracers))
mat
```

Let's check the proportions and total abundances by this scheme.

```{r}
scheme_df <- mat %>%
  as_tibble(rownames = 'sample') %>%
  pivot_longer(starts_with('tracer'), 
    names_to = 'tracer', values_to = 'dilution_name') %>%
  left_join(dilution_df, by = 'dilution_name') %>%
  with_groups(sample, mutate,
    proportion = conc / sum(conc)) %>%
  glimpse
```

```{r}
scheme_df %>%
  ggplot(aes(y = tracer, 
      x = proportion,
      #> x = proportion / (1 - proportion), 
      color = tracer)) +
  scale_x_log10() +
  ggbeeswarm::geom_quasirandom(groupOnX = FALSE)
```

```{r}
mat_conc <- scheme_df %>%
  metacal::build_matrix(sample, tracer, conc) %>%
  print(digits = 2)
mat_prop <- scheme_df %>%
  metacal::build_matrix(sample, tracer, proportion) %>%
  print(digits = 2)
```

a problem with this naive approach is that Tracer 2 is never in a large relative abundance.

Basic check: Each tracer is getting added in the same total amount:

```{r}
mat_conc %>% colSums
```

What is the range of total tracer in each sample?

```{r}
mat_conc %>% rowSums %>% log10
```


## output

```{r}
mat %>%
  as_tibble(rownames = 'sample') %>%
  write_csv('sample-layout.csv')
```

```{r}
mat %>%
  as_tibble(rownames = 'sample') %>%
  googlesheets4::write_sheet(
  ss = 'https://docs.google.com/spreadsheets/d/1kg_7zHF3O_IGj4Il9nCr-fZfmUdq0BcR9Y8wRyexaQ8/edit#gid=912041240', sheet = 1)
```

## Notes

Would potentially be nice to have the proportions above 1e-4 more evenly distributed.

Might also want to add sample 8 as a sample with none of the tracers spiked in.


## Sequencing

We need to pick the index barcodes to use for forward and reverse indexes.
Choose to be distinct from.

We can get the ONT barcodes from the seq barcode table,

```{r}
sb <- here('output/2022-05-18-sequencing-barcodes/sequencing-barcodes.csv') %>%
  read_csv
tracer_ont_ids <- str_c('BC', c(26, 29, 17, 60, 33))
fwd <- sb %>% slice(1) %>% pull(primer_1)
rev <- sb %>% slice(1) %>% pull(primer_2)
```

Let's reproduce Anjali's analysis in  https://docs.google.com/spreadsheets/d/193HrcFhwOGRD9WOa1g_Nc_k5rPGlxOcQcr30IFSbt6U/edit#gid=0

```{r}
ss <- 'https://docs.google.com/spreadsheets/d/193HrcFhwOGRD9WOa1g_Nc_k5rPGlxOcQcr30IFSbt6U/edit#gid=0'
order <- googlesheets4::read_sheet(ss)
```


```{r}
y <- sb %>%
  filter(! ont_barcode_id %in% tracer_ont_ids) %>%
  select(primer_1, primer_2, ont_barcode_id, ont_barcode_sequence) %>%
  pivot_longer(
    starts_with('primer'), values_to = 'pcr_primer'
  ) %>%
  mutate(
    full_primer = str_c('NN', ont_barcode_sequence, pcr_primer)
  ) %>%
  filter(
    (name == 'primer_1' & ont_barcode_id <= 'BC30') |
      (name == 'primer_2' & (ont_barcode_id > 'BC30' & ont_barcode_id <= 'BC63'))
  ) %>%
  glimpse
```

```{r}
y1 <- order %>%
  left_join(y, by = c('Barcode Name' = 'ont_barcode_id'))
all.equal(y1 %>% pull('Full Primer'), y1 %>% pull(full_primer))
```



